---
  title: "mixopt"
  output:
    html_document:
      toc: false
---

This documentation provides a brief introduction to mixopt. All functions can be imported from loading 'mixopt.R'. First of all, set working directory as a cloned repository on your local machine.

```{r}
# setwd("~/path/to/this/file");
# for instance, this file's directory is "~/git/mixopt/code" on my computer.
getwd();
source("mixopt.R");
```
  
The following code will install 'rjulia' package if you don't have it. You must have julia 0.5.2 for compatibility (Please see [rjulia repository][rjulia] for details). You may install julia package for the first time by setting julia_package_install = TRUE.

```{r}
mixopt_init(julia_package_install = FALSE);
```

Once installed and initialized, check if the following code works. Precompiling...

```{r}
out = mixopt_ex(100,2);
```

Let's try again...

```{r}
out = mixopt_ex(100000,1.1);
```

```{r}
plot(out$pihat,type='l',xlim = c(0,length(out$pihat)));
```

[rjulia]: https://github.com/armgong/rjulia